---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# uljas

<!-- badges: start -->
[![R-CMD-check](https://github.com/jhuovari/uljas/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuovari/uljas/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The R client for the Finnish Customs (Tulli) database Uljas. See:
https://tilastot.tulli.fi/en/uljas-statistical-database/uljas-api

## Installation

You can install the uljas from [github](https://github.com/jhuovari/uljas) with:

``` r
# install.packages("devtools")
devtools::install_github("jhuovari/uljas")
```

## Use

### Statistics

Statistics present in the database.

```{r stats}

library(uljas)

stats <- uljas_stats(lang = "en")

knitr::kable(stats)
```


### Dimensions

`uljas_dims` return dimensions that are available for a certain statistics. Use ifile from `uljas_stats`.

```{r dims}

sitc_dims <- uljas_dims(ifile = "/DATABASE/01 ULKOMAANKAUPPATILASTOT/02 SITC/ULJAS_SITC")

str(sitc_dims)
```


### Classifications

`uljas_class` returns values in the classification (specified with class parameter) for statistics (specified with ifile parameter). Also all (in main classifications) with `class = NULL`.

```{r class}

sitc_dims$`Classification of Products SITC`$label

sitc_class <- uljas_class(ifile = "/DATABASE/01 ULKOMAANKAUPPATILASTOT/02 SITC/ULJAS_SITC", class = "SITC Products")
#   sitc_class_all <- uljas_class(ifile = "/DATABASE/01 ULKOMAANKAUPPATILASTOT/02 SITC/ULJAS_SITC", class = NULL)

str(sitc_class)

```


### Data

`uljas_data` returns the data for class value combinations from a statistics (specified with ifile parameter). 

```{r data}

sitc_query <- list(`Classification of Products SITC1` = c("0" , "1"), `Time period` = "=ALL", Flow = 1, Country = "AT", Indicators = "V1")
sitc_data <- uljas_data(ifile = "/DATABASE/01 ULKOMAANKAUPPATILASTOT/02 SITC/ULJAS_SITC", classifiers = sitc_query)

head(sitc_data)

```

A classifier value may also be one of the api keywords `"=ALL"`, `"=FIRST"` and `"=LAST"`, the two latter optionally with a number of values, e.g. `"=LAST 12"`. The keywords follow the order of the classification, which for time periods is the newest first, so `"=FIRST"` is the latest period. The code of a total is an empty string, so a total can not be asked for on its own, but it is included in `"=ALL"`.

### Large queries

An individual request is limited to 50 000 cells, i.e. to 50 000 combinations of the classifier values. `uljas_query_size` tells how many cells a query asks for.

```{r size}

sitc5_query <- list(`Classification of Products SITC5` = "=ALL", `Time period` = "=ALL",
                    Flow = 1, Country = "AT", Indicators = "V1")

uljas_query_size(ifile = "/DATABASE/01 ULKOMAANKAUPPATILASTOT/02 SITC/ULJAS_SITC", classifiers = sitc5_query)

```

`uljas_data` divides a larger query into several requests and combines the answers, so the result is the same as that of a single request. Use `max_cells = Inf` to send the query as it is.

```{r large}

sitc5_data <- uljas_data(ifile = "/DATABASE/01 ULKOMAANKAUPPATILASTOT/02 SITC/ULJAS_SITC", classifiers = sitc5_query)

nrow(sitc5_data)

```

Note that the classifications and the statistics files of the database do
change. Finnish Customs renewed most of the cubes on 30 March 2026: the
current cubes (for example `ULJAS_SITC`) begin from January 2025 and the
longer time series are in the parallel cubes (for example `ULJAS_SITC2`).
Check the codes with `uljas_class` if a query stops returning data.
