---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# uljas

<!-- badges: start -->
<!-- badges: end -->

The R client for the Finnish Customs (Tulli) database Uljas. See: https://tulli.fi/en/statistics/uljas-api

## Installation

You can install the uljas from [github](https://github.com/jhuovari/uljas) with:

``` r
# install.packages("devtools")
devtools::install_github("jhuovari/uljas")
```

## Use

### Statistics

Statistics present in the database.

```{r stats}

library(uljas)

stats <- uljas_stats(lang = "en")

knitr::kable(stats)
```


### Dimensions

`uljas_dims` return dimensions that are available for a certain statistics. Use ifile from `uljas_stats`.

```{r dims}

sitc_dims <- uljas_dims(ifile = "/DATABASE/01 ULKOMAANKAUPPATILASTOT/02 SITC/ULJAS_SITC")

str(sitc_dims)
```


### Classifications

`sitc_class` returns values in the classification (specified with class parameter) for statistics (specified with ifile parameter). Also all (in main classifications) with `class = NULL`.

```{r class}

sitc_dims$`Classification of Products SITC`$label

sitc_class <- uljas_class(ifile = "/DATABASE/01 ULKOMAANKAUPPATILASTOT/02 SITC/ULJAS_SITC", class = "SITC Products")
#   sitc_class_all <- uljas_class(ifile = "/DATABASE/01 ULKOMAANKAUPPATILASTOT/02 SITC/ULJAS_SITC", class = NULL)

str(sitc_class)

```


### Data

`sitc_query` returns the data for class value combinations from a statistics (specified with ifile parameter). 

```{r data}

sitc_query <- list(`SITC Products` = c("0" , "02"), TimePeriod = "=ALL", Flow = 1, Country = "AT", Indicators = "V1")
sitc_data <- uljas_data(ifile = "/DATABASE/01 ULKOMAANKAUPPATILASTOT/02 SITC/ULJAS_SITC", classifiers = sitc_query)

head(sitc_data)

```



An individual request return is limited to 50 000 cells.

